Another Look at Regularized Discriminant Analysis for High-dimensional Classification

### 2013-02-05 reports/paper-etrda.tex
Added document containing alternative definition of RDA. Integrated the document into the latest draft of the paper.

\title{A Generalization of Regularized Discriminant Analysis with Applications to High-dimensional Microarray Data}

@@

\begin{abstract}
\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance matrix estimator, which is shrunken towards a scaled identity matrix. We demonstrate empirically, however, that the RDA classifier is often numerically incalculable or unstable for small-sample, high-dimensional data because the shrinkage factor is a function of the average of the eigenvalues, which tends to zero in this setting. Alternatively, \cite{Srivastava:2007ww} propose an empirical-Bayesian estimator that employs a similar shrinkage factor using the average of the nonzero eigenvalues.  Because the eigenvalues are often highly skewed, this average is often excessively large, so that the near-zero eigenvalues are substantially overcompensated. A comparable eigenvalue adjustment method from \cite{Thomaz:2006ef} often yields a shrinkage factor that is approximately equal to the average of the eigenvalues. By writing each shrinkage method as a matrix function, we generalize the shrinkage method of \cite{Thomaz:2006ef} to allow a powerful estimation technique that does not over inflate the near-zero eigenvalues. By incorporating our proposed covariance-matrix regularization with the RDA classifier, we overcome the numerical incalculability issues of the \emph{RDA} classifier. Moreover, we show that our proposed classifier is often superior to the regularized classifiers mentioned above in terms of classification accuracy using three simulated and five small-sample, high-dimensional microarray data sets. Additionally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
\end{abstract}

@@

(\textbf{Remove this paragraph?})
Despite its popularity, we demonstrate empirically that the covariance-matrix shrinkage employed in the \emph{RDA} classifier is often ineffective in adjusting the eigenvalues of the \emph{RDA} covariance-matrix estimator when $p \gg n$. The \emph{RDA} classifier is numerically unstable as the number of features incorporated into the \emph{RDA} classifier is increased because the covariance-matrix estimator is nearly singular. Specifically, the shrinkage applied in the \emph{RDA} classifier is a function of the average eigenvalues of the convex combination of the class sample covariance matrix and the sample covariance matrix pooled across all classes. Because the average of the eigenvalues is often nearly zero when $p \gg n$, the resulting covariance-matrix estimator has two issues. First, the determinant of the covariance-matrix estimator approaches zero, suggesting that the \emph{RDA} covariance-matrix estimator is nearly singular in the small-sample, high-dimensional setting. Second, the smallest eigenvalues are near zero, which causes the \emph{RDA} classifier to have an incalculable or numerically unstable discriminant function.

@@

Using \textbf{five} small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers. We also include in our comparison the alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of diagonal discriminant analysis from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. Additionally, we examine the classification accuracy of the classifiers considered using \textbf{three} simulation configurations. Furthermore, we show that our proposed eigenvalue-threshold methods are more numerically stable than the other considered regularization methods using the simulated configurations.

We have organized the remainder of the paper as follows. In Section 2 we review the \emph{RDA}, the \emph{MDEB}, and \emph{NLDA} classifiers and define each in terms of matrix functions. In Section 3 we generalize the eigenvalue adjustment employed in the \emph{NLDA} classifier and define our proposed \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers. In Section 4 we describe the \textbf{three} simulation configurations and \textbf{five} small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 6.

\section{Covariance-Matrix Regularization in Discriminant Analysis}

In discriminant analysis we wish to correctly assign an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \mathcal{K} = \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{m \times n}$ denotes the matrix space of all $m \times n$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ is the MLE for $\bm \mu_k$ and  $\widehat{\bm\Sigma}_k$ is the MLE for $\bm \Sigma_k$. Notice that if we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
\begin{align}
	D_{LDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}^{-1}(\bm x - \xbar_k), \label{eq:lda}
\end{align}
where $\widehat{\bm\Sigma}$ is the pooled sample covariance matrix
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}

In the small-sample, high-dimensional setting, if $p > n_k$ for any $k$, \eqref{eq:qda} is incalculable because the sample covariance matrix $\widehat{\bm\Sigma}_k$ for class $\omega_k$ is singular, and, thus, its inverse $\widehat{\bm\Sigma}_k^{-1}$ does not exist.

Although \eqref{eq:qda} is in a standard form, we prefer to express the \emph{QDA} classifier explicitly in terms of the eigenvalues and eigenvectors of $\widehat{\bm \Sigma}_k$. Let $e_{jk}$ denote the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k$ and $\bm v_{jk}$ be the corresponding eigenvector such that $e_{1k} \ge e_{2k} \ge \ldots \ge e_{pk} \ge 0$. Equivalently, we write \eqref{eq:qda} as
\begin{align}
  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \frac{[\bm v_{jk}' (\bm x - \bm \xbar_k)]^2}{e_{jk}} + \sum_{j=1}^p \ln e_{jk}. \label{eq:qda-eigen}
\end{align}
The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda-eigen} . In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}. This bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that the last $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0, which implies that \eqref{eq:qda-eigen} is incalculable. Thus, although more feature information is available to discriminate among the $K$ classes as $p$ increases, classification accuracy decreases unless one obtains enough training-sample observations to reliably estimate the increased number of parameters.

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to stabilize the eigenvalues of $\widehat{\bm \Sigma}_k$ used in \eqref{eq:qda}. These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}. Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$. Then
\begin{align}
	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
\end{align}
is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$. For this reason, $\phi$ is often called a spectral transform. Special case of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.

Viewing regularized covariance-matrix estimators as matrix functions, we can write a more general form of the \emph{QDA} classifier as
\begin{align}
  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \phi(e_{jk})[\bm v_{jk}' (\bm x - \bm \mu_k)]^2 + \sum_{j=1}^p \ln \{ \phi(e_{jk}) \}. \label{eq:generalized-qda}
\end{align}
The generalized \emph{QDA} classifier in \eqref{eq:generalized-qda} includes several common regularized classifiers. Notice that if  $\phi(e_{jk}) = e_{jk}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{QDA} classifier in \eqref{eq:qda-eigen}. Furthermore, if $\phi(e_{jk}) = e_{j}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{LDA} classifier in \eqref{eq:lda}, and the second summand in \eqref{eq:generalized-qda} is constant across all classes and can be omitted. Additionally, one common approach is to ignore the zero eigenvalues when $p > n$ by employing the Moore-Penrose pseudoinverse $\widehat{\bm \Sigma}_k^{+}$, which employs the matrix function $\phi(e_j) = e_j^{+}$, where $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise. As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.

Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\phi(e_{jk}) = (e_{jk} + \gamma_k)^{-1}, \label{eq:ridge-estimator}
\end{align}
is a ridge-like estimator, similar to a method employed in ridge regression. The estimator in \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$ so that they are positive, resulting in $\widehat{\bm \Sigma}_k(\gamma_k)$ being positive definite.

@@

\subsection{Regularized Discriminant Analysis}

\cite{Friedman:1989tm} has extended the shrinkage technique by first computing a
weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class
$\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to
estimate the covariance matrix for class $\omega_k$ with $\widehat{\bm\Sigma}_k(\lambda)$,
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$, and
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda).\label{eq:sig-lambda}
\end{align}
We can interpret \eqref{eq:sig-lambda} as a
covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. Notice that for $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Similarly, notice that for $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.

>>>>

\title{Another Look at Regularized Discriminant Analysis for High-dimensional Classification}

@@

\begin{abstract}
\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis (\emph{RDA}) classifier for the small-sample, high-dimensional setting by employing a regularized covariance matrix estimator, which employs a ridge-like shrinkage factor to ensure positive definiteness. We demonstrate that the \emph{RDA} classifier, in its original form, is inadequate for the small-sample, high-dimensional setting. We redefine the \emph{RDA} classifier to increase its computational runtime and its interpretation. We show that this new definition includes several well-known regularized classifiers as special cases. We also consider several properties of the newly defined \emph{RDA} classifier. Alternatively, \cite{Srivastava:2007ww} propose an empirical-Bayesian estimator that employs a similar shrinkage factor using the average of the nonzero eigenvalues.  Because the eigenvalues are often highly skewed, this average is often excessively large, so that the near-zero eigenvalues are substantially overcompensated. A comparable eigenvalue adjustment method from \cite{Thomaz:2006ef} often yields a shrinkage factor that is approximately equal to the average of the eigenvalues. By writing each shrinkage method as a matrix function, we generalize the shrinkage method of \cite{Thomaz:2006ef} to allow a powerful estimation technique that does not over inflate the near-zero eigenvalues. By incorporating our proposed covariance-matrix regularization with the RDA classifier, we overcome the numerical incalculability issues of the \emph{RDA} classifier. Moreover, we show that our proposed classifier is often superior to the regularized classifiers mentioned above in terms of classification accuracy using three simulated and three small-sample, high-dimensional microarray data sets. Additionally, we provide implementations of the competing classifiers in the {\tt regdiscrim} R package, available on CRAN.
\end{abstract}

@@

@@

Using three small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers. We also include in our comparison the alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of diagonal discriminant analysis from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. Additionally, we examine the classification accuracy of the classifiers considered using \textbf{three} simulation configurations. Furthermore, we show that our proposed eigenvalue-threshold methods are more numerically stable than the other considered regularization methods using the simulated configurations.

We have organized the remainder of the paper as follows. In Section 2 we review the \emph{RDA}, the \emph{MDEB}, and \emph{NLDA} classifiers and define each in terms of matrix functions. In Section 3 we generalize the eigenvalue adjustment employed in the \emph{NLDA} classifier and define our proposed \emph{Soft-ETRDA} and \emph{Hard-ETRDA} classifiers. In Section 4 we describe the \textbf{three} simulation configurations and three small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 6.

\section{Regularized Discriminant Analysis}

We\footnote{Define the majority of notation here, including $z^+$, $z_+$, and $I[z]$.} wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \mathcal{K} = \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{m \times n}$ denotes the matrix space of all $m \times n$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier,
\begin{align}
	D_{LDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}^{-1}(\bm x - \xbar_k), \label{eq:lda}
\end{align}
where $\widehat{\bm\Sigma}$ is the pooled sample covariance matrix
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}

\subsection{Original Definition of Regularized Discriminant Analysis}

\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}$ used in the \emph{LDA} classifier in \eqref{eq:lda}, in which case we assume implicitly that $\bm\Sigma_k = \bm \Sigma$ for $k = 1, \ldots, K$.

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased \emph{RDA} covariance-matrix estimator
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
\end{align}
for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter that controls the shrinkage of \eqref{eq:sig-rda}  towards the $p$-dimensional identity matrix $\bm I_p$, which is weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from $\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the amount of shrinkage applied.

Substituting \eqref{eq:sig-rda} into \eqref{eq:qda}, we have the \emph{RDA} classifier, where we assign an unlabeled observation $\bm x$ to class $\omega_k$:
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}

\subsection{An Alternative Approach to Regularized Discriminant Analysis}

Here, we present a more straightforward, alternative parameterization of the \emph{RDA} classifier. We first define explicitly
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. Before examining the shrinkage applied to \eqref{eq:sig-lambda-alternative}, we examine some of its properties and write \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'}.\label{eq:sig-lambda-alternative2}
\end{align}
From \eqref{eq:sig-lambda-alternative2}, we see the weight $\lambda$ places on the covariance matrices from the remaining $K - 1$ classes. The interesting part here is that $\lambda$ controls the contribution of each of the $N$ observations to estimating $\Sigma_k$. Effectively, this states that for $0 < \lambda \le 1$, we use $N$ observations to estimate $\bm \Sigma_k$ rather than $n_k$ as done in the \emph{QDA} classifier. As we will discuss in-depth later, the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from $n_k - 1$ to $N - 1$ for $0 < \lambda \le 1$. Moreover, suppose that we have centered each observation $\bm x_i$, $i = 1, \ldots, N$, by its class sample mean. Then, we write \eqref{eq:sig-lambda-alternative2} as
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative3}
\intertext{where}
	c_{ik}(\lambda) &= \begin{cases}
		(1 - \lambda)n_k^{-1} + \lambda N^{-1}, & y_i = k,\\
		\lambda N^{-1}, & y_i \ne k.\\
	\end{cases}\nonumber
\end{align}
Therefore, after centering each observation by its corresponding class sample mean, we have that $\widehat{\bm\Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda-alternative3} incorporates each centered observation into the estimation of $\bm \Sigma_k$. Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative3} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.

\subsection{Covariance Matrix Regularization}

Although \eqref{eq:qda} is in a standard form, we prefer to express the \emph{QDA} classifier explicitly in terms of the eigenvalues and eigenvectors of $\widehat{\bm \Sigma}_k$. Let $e_{jk}(\lambda)$ denote the $j$th largest eigenvalue of $\widehat{\bm\Sigma}_k$ and $\bm v_{jk}(\lambda)$ be the corresponding eigenvector such that $e_{1k}(\lambda) \ge e_{2k}(\lambda) \ge \ldots \ge e_{pk}(\lambda) \ge 0$. Equivalently, we write \eqref{eq:qda} as
\begin{align}
  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \frac{[\bm v_{jk}(\lambda)' (\bm x - \bm \xbar_k)]^2}{e_{jk}(\lambda)} + \sum_{j=1}^p \ln \{e_{jk}(\lambda)\}. \label{eq:rda-eigen}
\end{align}
The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:rda-eigen}. In fact, the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}. This bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k(\lambda)) \le n_k$, which implies that the last $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ are 0, which implies that \eqref{eq:rda-eigen} is incalculable. Thus, although more feature information is available to discriminate among the $K$ classes as $p$ increases, classification accuracy decreases unless one obtains enough training-sample observations to reliably estimate the increased number of parameters.

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to stabilize the eigenvalues of $\widehat{\bm \Sigma}_k$ used in \eqref{eq:qda}. These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}. Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$. Then
\begin{align}
	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
\end{align}
is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$. For this reason, $\phi$ is often called a spectral transform \citep{Ch15SSLBook} or a transfer function \citep{Ye:2009gd, Chapelle2002SSLChapter15Reference} . Special case of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.

Viewing regularized covariance-matrix estimators as matrix functions, we can write a more general form of the \emph{QDA} classifier as
\begin{align}
  D(\bm x) = \argmin_{k} \sum_{j = 1}^p \phi(e_{jk}(\lambda))[\bm v_{jk}(\lambda)' (\bm x - \bm \mu_k)]^2 + \sum_{j=1}^p \ln \{ \phi(e_{jk}(\lambda)) \}. \label{eq:generalized-qda}
\end{align}
The\footnote{In \eqref{eq:generalized-qda} we have a slight issue in the formulation of the log-determinant in the second sum. The eigenvalue function $\phi$ in the first sum should be the inverse of the function in the second sum. Furthermore, we need to define the log-determinant so that the last $p - q$ terms are ignored when we reduce the dimension to $q$.} generalized \emph{QDA} classifier in \eqref{eq:generalized-qda} includes several common regularized classifiers. Notice that if  $\phi(e_{jk}(\lambda)) = e_{jk}(\lambda)$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{QDA} classifier in \eqref{eq:rda-eigen}. Furthermore, if $\phi(e_{jk}(\lambda)) = e_{j}$, $k = 1, \ldots, K$, \eqref{eq:generalized-qda} reduces to the \emph{LDA} classifier in \eqref{eq:lda}, and the second summand in \eqref{eq:generalized-qda} is constant across all classes and can be omitted. Additionally, one common approach is to ignore the zero eigenvalues when $p > n$ by employing the Moore-Penrose pseudoinverse $\widehat{\bm \Sigma}_k^{+}$, which employs the matrix function $\phi(e_j) = e_j^{+}$, where $z^{+} = 1/z$ if $z > 0$ and $0$ otherwise. As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.

Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\phi(e_{jk}(\lambda)) = (e_{jk}(\lambda) + \gamma_k)^{-1}, \label{eq:ridge-estimator}
\end{align}
is a ridge-like estimator, similar to a method employed in ridge regression. The estimator in \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k(\lambda)$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$ by $\gamma_k$ so that they are positive, resulting in $\widehat{\bm \Sigma}_k(\lambda) + \gamma_k \bm I_p$ being positive definite.

@@

\section{Properties of the Redefined RDA Classifier}

\subsection{Computational Advantages}

Thus far, our alternative parameterization simplifies the definition of the \emph{RDA} classifier and has an improved interpretation in terms of the contribution of each observation. Here, we demonstrate that our parameterization yields significant computational advantages over the originally proposed \emph{RDA} classifier from \cite{Friedman:1989tm} by writing \eqref{eq:sig-lambda-alternative3} as
\begin{align*}
	\widehat{\bm\Sigma}_k(\lambda) &= \bm X_k(\lambda)' \bm X_k(\lambda),
\end{align*}
where $\bm X_k(\lambda) = [\sqrt{c_{1k}(\lambda)} \bm x_1', \ldots,  \sqrt{c_{Nk}(\lambda)} \bm x_N']'$. Now, we can apply the Fast SVD to compute the eigenvalue decomposition of $\widehat{\bm\Sigma}_k(\lambda)$ efficiently. By writing the SVD of $\bm X_k(\lambda) = \bm U_k \bm D_k \bm V_k'$, we have the spectral decomposition $\widehat{\bm\Sigma}_k(\lambda) = \bm X_k(\lambda)' \bm X_k(\lambda) = \bm V_k \bm D_k^2 \bm V_k'$, where $\bm U_k \in \mathbb{R}_{N \times N}$ is orthogonal, $\bm D_k \in \mathbb{R}_{N \times N}$ is a diagonal matrix consisting of the singular values of $\bm X_k(\lambda)$, and $\bm V_k \in \mathbb{R}_{p \times N}$ is orthogonal. Notice that $\bm X_k(\lambda)\bm X_k(\lambda)' = \bm U_k \bm D_k^2 \bm U_k'$, so that the eigenvalue decomposition of $\bm X_k(\lambda)\bm X_k(\lambda)'$ yields $\bm U_k$ and $\bm D_k$. This is much faster when $p \gg N$ because $\bm X_k(\lambda)\bm X_k(\lambda)' \in \mathbb{R}_{N \times N}$. Also, notice that the singular values of $\bm D_k$ are nonzero, so that $\bm D_k^{-1}$ exists. After obtaining $\bm U_k$ and $\bm D_k$, we compute the matrix consisting of the first $N$ eigenvectors of $\widehat{\bm\Sigma}_k(\lambda)$ by simply computing\footnote{TODO: Update this paragraph to theorem/proof.}
\begin{align*}
	\bm V_k = \bm X_k(\lambda)' \bm U_k \bm D_k^{-1}.
\end{align*}

\subsection{Covariance-Matrix Regularization}
